On exponential sums arising from the classical doubling method

### 2025-03-25 reorganization.tex
Calvin changes

\subsubsection{Rational Tori}
A \emph{rational torus} $\algebraicGroup{T} \subseteq \algebraicGroup{G}$ over $\finiteField$ is a torus defined over $\finiteField$. 
Over a finite field, $\algebraicGroup{G}(\finiteField)$ conjugacy classes of maximal rank rational tori \calvin{check if we need connected or something} are parameterized by  Frobenius order twisted conjugacy classes in the Weyl group. In particular if you fix a rational torus $\algebraicGroup{T}_0$ with rank $=\mathrm{rk} \algebraicGroup{G}$ and an element $g \in \algebraicGroup{G}(\algebraicClosure{\finiteField})$ such that $g^{-1}\Frobenius(g) \in N(\algebraicGroup{T}_0)$ then the torus $g\algebraicGroup{T}_0g^{-1}$ is rational, and it's $\algebraicGroup{G}(\finiteField)$-conjugacy class is determined  by the Frobeinus-twisted conjugacy class of the image of $g^{-1}\Frobenius(g)$ in $W(\algebraicGroup{T}_0)(\algebraicClosure{\mathbb{F}})$, see \cite[Section 3.3]{Carter1985}.

%since $\Frobenius(g\algebraicGroup{T}_0g^{-1}) = g\algebraicGroup{T}_0g^{-1}$
%could really demonstrate this by stating Land map is surjective + that all maximal tori are conjugate over an algebraiclly closed field
Applying this theory to our groups $\algebraicGroup{G}$, for each group we find a nice rational torus $\algebraicGroup{T}$ and write out the Frobenius order action on the Weyl group and the

\begin{enumerate}
	\item(Special orthogonal groups) 
	In each case let $\algebraicGroup{T}$ be the diagonal torus. Then $W(\algebraicGroup{T} )$ has representatives the type B or D permutation matrices which are fixed by the Frobenius order.
	\begin{enumerate}[(a)]
		\item (Split odd case $\algebraicGroup{SO}_{2n+1}$) $\algebraicGroup{T} = \algebraicGroup{G}_m^{n}$. In type B the Weyl group conjugacy classes are given by $(\lambda^+,\lambda^-)$ where $\lambda^+$ is the partition formed from and $\lambda^-$ is the partition formed from. Given a representation $w$ of the conjugacy class we get:
		$$^w\algebraicGroup{T} = \prod_{i} \restrictionOfScalars{\finiteFieldExtension{\lambda^+_i}}{\finiteField}{\multiplcativeScheme} \times \prod_{i} \restrictionOfScalars{\finiteFieldExtension{\lambda^-_i}}{\finiteField}{\algebraicGroup{\UnitaryGroup}_1} $$
		\item (Split even case $\algebraicGroup{SO}_{2n}^+$) $\algebraicGroup{T} = \algebraicGroup{G}_m^{n}$.  Given $w \in W(\algebraicGroup{T})$ the rational structure of $^w\algebraicGroup{T}$ is determined by the conjugacy class of $w$ in the Type B Weyl group (a $\mathbb{Z}/2\mathbb{Z}$ extension of $W(\algebraicGroup{T})$. If $w$ is conjugate in the extension to $(\lambda^+,\lambda^-)$ where $\lambda^+$ is the partition formed from and $\lambda^-$ is the partition formed from:
		$$^w\algebraicGroup{T} = \prod_{i} \restrictionOfScalars{\finiteFieldExtension{\lambda^+_i}}{\finiteField}{\multiplcativeScheme} \times \prod_{i} \restrictionOfScalars{\finiteFieldExtension{\lambda^-_i}}{\finiteField}{\algebraicGroup{\UnitaryGroup}_1} $$
			\item (Non-split even case $\algebraicGroup{SO}_{2n}^-$) Set $\algebraicGroup{T} = \algebraicGroup{G}_m^{n-1} \times \algebraicGroup{U}(1)$ the diagonal torus. Given $w \in W(\algebraicGroup{T})$ the rational structure of $^w\algebraicGroup{T}$ is determined by the conjugacy class of $w$ in the Type B Weyl group (a $\mathbb{Z}/2\mathbb{Z}$ extension of $W(\algebraicGroup{T})$. If $w \circ (-1,1)$ is conjugate in the extension to $(\lambda^+,\lambda^-)$ where $\lambda^+$ is the partition formed from and $\lambda^-$ is the partition formed from:
		$$^w\algebraicGroup{T} = \prod_{i} \restrictionOfScalars{\finiteFieldExtension{\lambda^+_i}}{\finiteField}{\multiplcativeScheme} \times \prod_{i} \restrictionOfScalars{\finiteFieldExtension{\lambda^-_i}}{\finiteField}{\algebraicGroup{\UnitaryGroup}_1} $$
	\end{enumerate}
	\item (Symplectic groups) In type B the Weyl group conjugacy classes are given by $(\lambda^+,\lambda^-)$ where $\lambda^+$ is the partition formed from and $\lambda^-$ is the partition formed from. Given a representation $w$ of the conjugacy class we get:
		$$^w\algebraicGroup{T} = \prod_{i} \restrictionOfScalars{\finiteFieldExtension{\lambda^+_i}}{\finiteField}{\multiplcativeScheme} \times \prod_{i} \restrictionOfScalars{\finiteFieldExtension{\lambda^-_i}}{\finiteField}{\algebraicGroup{\UnitaryGroup}_1} $$
	\item (Unitary groups) We let our torus be the diagonal torus for the unitary group with hermitian form $I_n$. Then $\algebraicGroup{T}=\algebraicGroup{\UnitaryGroup}_1^n$ and the Frobenius action on the Weyl group is trivial. Note since all hermitian forms are equivalent over a finite field, $\algebraicGroup{T}$ is a torus in the group $\algebraicGroup{U}_n$ from the previous section.  In type A the Weyl group conjugacy classes are given by $(\lambda)$, and
	$$^w\algebraicGroup{T} = \prod_{i} \begin{dcases}
	\restrictionOfScalars{\finiteFieldExtension{\lambda^+_i}}{\finiteField}{\multiplcativeScheme} & \lambda_i \textnormal{ is even} \\
	\restrictionOfScalars{\finiteFieldExtension{\lambda^+_i}}{\finiteField}{\algebraicGroup{\UnitaryGroup}_1}  & \lambda_i \textnormal{ is odd}
	\end{dcases}$$
	\item (Even special orthogonal similitude groups)
	\item (Symplectic similitude groups)
\end{enumerate}

@@

Let $\algebraicGroup{A}_n \subset \algebraicGroup{\GL}_n$ be the diagonal torus. We denote for an element $g \in \algebraicGroup{\GL}_n$, $g^{\ast} = w_n \transpose{g}^{-1} w_n$, where $w_n$ is the long Weyl element (see \Cref{subsec:isometry-groups}). Using $\algebraicGroup{A}_n$ we define standard (mostly split) tori in our groups as follows.
\begin{center}
\begin{tabular}{|c|c|}
	\hline 
		$\algebraicGroup{G}$ & $\algebraicGroup{T}^{\mathrm{std}}_{\algebraicGroup{G}}$ \tabularnewline
		\hline 
		\hline 
		$\algebraicGroup{\UnitaryGroup}_{2n}$ & $\left\{ \begin{pmatrix}
			a\\
			& \involution{\left(a^{\ast}\right)}
		\end{pmatrix}  \mid a \in \algebraicGroup{A}_n \right\}$ \tabularnewline
		\hline 
		$\algebraicGroup{\UnitaryGroup}_{2n+1}$ & $\left\{ \begin{pmatrix}
	a \\
	& t\\
	& & \involution{\left(a^{\ast}\right)}
\end{pmatrix}  \mid a \in \algebraicGroup{A}_n,\,\, t \in \algebraicGroup{U}_1 \right\}$ \tabularnewline
\hline 
	$\algebraicGroup{\SO}_{2n+1}$ & $\left\{ \begin{pmatrix}
		a\\
		& 1\\
		& & a^{\ast}
	\end{pmatrix} \mid a \in \algebraicGroup{A}_n \right\}$ \tabularnewline
	\hline 
	$\algebraicGroup{\Sp}_{2n}$ or $\algebraicGroup{\SO}^{+}_{2n}$ & $\left\{ \begin{pmatrix}
		a\\
		& a^{\ast}
	\end{pmatrix} \mid a \in \algebraicGroup{A}_n \right\}$ \tabularnewline
	\hline 
	$\algebraicGroup{\SO}_{2n}^{-}$ & $\left\{ \begin{pmatrix}
		a\\
		& h\\
		& & a^{\ast}
	\end{pmatrix} \mid a \in \algebraicGroup{A}_{n-1}\,\, h \in \algebraicGroup{\SO}_2^{-} \right\}$ \tabularnewline
	\hline		
	\end{tabular}
\end{center}
Here $\algebraicGroup{T}^{\mathrm{std}}_{\algebraicGroup{G}}$ is split for all rows except for the the last row.
Here $\algebraicGroup{T}^{\mathrm{std}}_{\algebraicGroup{G}}$, is split for all rows except for the the last row. We also write down the standard tori for the relevant corresponding similitude groups.
\begin{center}
	\begin{tabular}{|c|c|}
		\hline 
		$\algebraicGroup{G}$ & $\algebraicGroup{T}^{\mathrm{std}}_{\algebraicGroup{G}}$ \tabularnewline
		\hline 
		\hline 
		$\algebraicGroup{\GSp}_{2n}$ or $\algebraicGroup{\GSO}^{+}_{2n}$ & $\left\{\left(\begin{pmatrix}
			\lambda a\\
			& a^{\ast}
		\end{pmatrix}, \lambda\right) \mid a \in \algebraicGroup{A}_n,\,\, \lambda \in \multiplcativeScheme \right\}$ \tabularnewline
		\hline 
		$\algebraicGroup{\GSO}_{2n}^{-}$ & $\left\{ \left(\begin{pmatrix}
			\algebraicGroup{\aFieldNorm}_{\finiteFieldExtension{2} \slash \finiteField}\left(h\right) a\\
			& h\\
			& & a^{\ast}
		\end{pmatrix}, \algebraicGroup{\aFieldNorm}_{\finiteFieldExtension{2} \slash \finiteField}\left(h\right)\right) \mid a \in \algebraicGroup{A}_{n-1}\,\, h \in \restrictionOfScalars{\finiteFieldExtension{2}}{\finiteField}{\multiplcativeScheme} \right\}$ \tabularnewline
		\hline		
	\end{tabular}
\end{center}
Here we realize $\restrictionOfScalars{\finiteFieldExtension{2}}{\finiteField}{\multiplcativeScheme}$ as $$\restrictionOfScalars{\finiteFieldExtension{2}}{\finiteField}{\multiplcativeScheme} = \left\{ \left(\begin{pmatrix}
	a & b\\
	db & a
\end{pmatrix}, \lambda\right) \mid a^2 - db^2 = \lambda \in \multiplcativeScheme \right\},$$
where $d$ is as in \Cref{subsec:isometry-groups}.

>>>>

\subsubsection{Rational Tori}
A \emph{rational torus} $\algebraicGroup{T} \subset \algebraicGroup{G}$ over $\finiteField$ is a torus defined over $\finiteField$. 
Over a finite field, $\algebraicGroup{G}(\finiteField)$ conjugacy classes of maximal rank rational tori \calvin{check if we need connected or something} are parameterized by  Frobenius order twisted conjugacy classes in the Weyl group. In particular if you fix a rational torus $\algebraicGroup{T}_0$ with rank $=\mathrm{rk} \algebraicGroup{G}$ and an element $g \in \algebraicGroup{G}(\algebraicClosure{\finiteField})$ such that $g^{-1}\Frobenius(g) \in N(\algebraicGroup{T}_0)$ then the torus $g\algebraicGroup{T}_0g^{-1}$ is rational, and it's $\algebraicGroup{G}(\finiteField)$-conjugacy class is determined  by the Frobeinus-twisted conjugacy class of the image of $g^{-1}\Frobenius(g)$ in $W(\algebraicGroup{T}_0)(\algebraicClosure{\mathbb{F}})$, see \cite[Section 3.3]{Carter1985}.

@@

Let $\algebraicGroup{A}_n \subset \algebraicGroup{\GL}_n$ be the diagonal torus. We denote for an element $g \in \algebraicGroup{\GL}_n$, $g^{\ast} = w_n \transpose{g}^{-1} w_n$, where $w_n$ is the long Weyl element (see \Cref{subsec:isometry-groups}). Using $\algebraicGroup{A}_n$ we define standard (mostly split) tori in our groups as follows.
\begin{center}
\begin{tabular}{|c|c|}
	\hline 
		$\algebraicGroup{G}$ & $\algebraicGroup{T}^{\mathrm{std}}_{\algebraicGroup{G}}$ \tabularnewline
		\hline 
		\hline 
		$\algebraicGroup{\UnitaryGroup}_{2n}$ & $\left\{ \begin{pmatrix}
			a\\
			& \involution{\left(a^{\ast}\right)}
		\end{pmatrix}  \mid a \in \algebraicGroup{A}_n \right\}$ \tabularnewline
		\hline 
		$\algebraicGroup{\UnitaryGroup}_{2n+1}$ & $\left\{ \begin{pmatrix}
	a \\
	& t\\
	& & \involution{\left(a^{\ast}\right)}
\end{pmatrix}  \mid a \in \algebraicGroup{A}_n,\,\, t \in \algebraicGroup{U}_1 \right\}$ \tabularnewline
\hline 
	$\algebraicGroup{\SO}_{2n+1}$ & $\left\{ \begin{pmatrix}
		a\\
		& 1\\
		& & a^{\ast}
	\end{pmatrix} \mid a \in \algebraicGroup{A}_n \right\}$ \tabularnewline
	\hline 
	$\algebraicGroup{\Sp}_{2n}$ or $\algebraicGroup{\SO}^{+}_{2n}$ & $\left\{ \begin{pmatrix}
		a\\
		& a^{\ast}
	\end{pmatrix} \mid a \in \algebraicGroup{A}_n \right\}$ \tabularnewline
	\hline 
	$\algebraicGroup{\SO}_{2n}^{-}$ & $\left\{ \begin{pmatrix}
		a\\
		& h\\
		& & a^{\ast}
	\end{pmatrix} \mid a \in \algebraicGroup{A}_{n-1}\,\, h \in \algebraicGroup{\SO}_2^{-} \right\}$ \tabularnewline
	\hline		
	\end{tabular}
\end{center}
Here, $\algebraicGroup{T}^{\mathrm{std}}_{\algebraicGroup{G}}$ is split for all rows except for the last row. We also write down the standard tori for the relevant corresponding similitude groups.
\begin{center}
	\begin{tabular}{|c|c|}
		\hline 
		$\algebraicGroup{G}$ & $\algebraicGroup{T}^{\mathrm{std}}_{\algebraicGroup{G}}$ \tabularnewline
		\hline 
		\hline 
		$\algebraicGroup{\GSp}_{2n}$ or $\algebraicGroup{\GSO}^{+}_{2n}$ & $\left\{\left(\begin{pmatrix}
			\lambda a\\
			& a^{\ast}
		\end{pmatrix}, \lambda\right) \mid a \in \algebraicGroup{A}_n,\,\, \lambda \in \multiplcativeScheme \right\}$ \tabularnewline
		\hline 
		$\algebraicGroup{\GSO}_{2n}^{-}$ & $\left\{ \left(\begin{pmatrix}
			\algebraicGroup{\aFieldNorm}_{\finiteFieldExtension{2} \slash \finiteField}\left(h\right) a\\
			& h\\
			& & a^{\ast}
		\end{pmatrix}, \algebraicGroup{\aFieldNorm}_{\finiteFieldExtension{2} \slash \finiteField}\left(h\right)\right) \mid a \in \algebraicGroup{A}_{n-1}\,\, h \in \restrictionOfScalars{\finiteFieldExtension{2}}{\finiteField}{\multiplcativeScheme} \right\}$ \tabularnewline
		\hline		
	\end{tabular}
\end{center}
Here we realize $\restrictionOfScalars{\finiteFieldExtension{2}}{\finiteField}{\multiplcativeScheme}$ as $$\restrictionOfScalars{\finiteFieldExtension{2}}{\finiteField}{\multiplcativeScheme} = \left\{ \left(\begin{pmatrix}
	a & b\\
	db & a
\end{pmatrix}, \lambda\right) \mid a^2 - db^2 = \lambda \in \multiplcativeScheme \right\},$$
where $d$ is as in \Cref{subsec:isometry-groups}.

In the unitary case, the action of $\Frobenius$ on $W(\algebraicGroup{T}_{\algebraicGroup{U}}^{std})$ is conjugation by $w_n$. In the $\algebraicGroup{SO}^-_{2n}$ case, the action of $\Frobenius$ on $W(\algebraicGroup{T}_{\algebraicGroup{U}}^{std})$ is the outer automorphism on the Dynkin diagram of type D. In terms of signed permutations with an even number of negative signs this corresponds to conjugation by the transposition $(-1,1)$ \calvin{or the $(-n,n)$. Kinda depends on conventions. The most correct thing to say is it corresponds to conjugation by the matrix $\begin{pmatrix} I_{n-1} & 0 & 0&0 \\ 0 &0&1&0\\0&1&0&0\\0&0&0&I_{n-1}\end{pmatrix}$}. In all other cases, the action of $\Frobenius$ on $W(\algebraicGroup{T}_{\algebraicGroup{U}}^{std})$ is trivial. We now parameterize the Weyl group $\Frobenius$-twisted conjugacy classes and their corresponding tori:

\begin{enumerate}
    \item (Type A: Unitary groups) The map $w \mapsto w\cdot w_n$ sends a $w_n$-twisted $S_n$ conjugacy class to an $S_n$ conjugacy class. Thus twisted conjugacy classes are determined by the cycle partition $\lambda$ associated to $w\cdot w_n$. The corresponding algebraic torus is \calvin{$L$ is the lang map}
	$$^{L^{-1}w}\algebraicGroup{T} =\prod_{i} \begin{dcases}
	\restrictionOfScalars{\finiteFieldExtension{\lambda^+_i}}{\finiteField}{\multiplcativeScheme} & \lambda_i \textnormal{ is even} \\
	\restrictionOfScalars{\finiteFieldExtension{\lambda^+_i}}{\finiteField}{\algebraicGroup{\UnitaryGroup}_1}  & \lambda_i \textnormal{ is odd}
	\end{dcases}$$
    \item (Type B, C, D: Special orthogonal and symplectic groups)
    Conjugacy classes in the type B Weyl (signed permutations) group are classified by pairs of partitions $(\lambda^+,\lambda^-)$.
	\begin{enumerate}[(a)]
        \item ($\algebraicGroup{SO}_{2n+1}$ or $\algebraicGroup{Sp}$) Let $(\lambda^+,\lambda^-)$ correspond to the conjugacy class of $w$. Then
		$$^{L^{-1}w}\algebraicGroup{T} = \prod_{i} \restrictionOfScalars{\finiteFieldExtension{\lambda^+_i}}{\finiteField}{\multiplcativeScheme} \times \prod_{i} \restrictionOfScalars{\finiteFieldExtension{\lambda^-_i}}{\finiteField}{\algebraicGroup{\UnitaryGroup}_1} $$
        \item ($\algebraicGroup{SO}_{2n}^+$) Viewing the type D Weyl (even signed permutations) group as a subgroup of signed permutations, we can associate to a $(\lambda^+,\lambda^-)$ pair via the conjugacy class of $w$. Then
		$$^{L^{-1}w}\algebraicGroup{T} = \prod_{i} \restrictionOfScalars{\finiteFieldExtension{\lambda^+_i}}{\finiteField}{\multiplcativeScheme} \times \prod_{i} \restrictionOfScalars{\finiteFieldExtension{\lambda^-_i}}{\finiteField}{\algebraicGroup{\UnitaryGroup}_1} $$
		\item ($\algebraicGroup{SO}_{2n}^-$) The map $w \mapsto w\circ (-1,1)$ sends a $(-1,1)$-twisted conjugacy class in the even signed permutation group to a conjugacy class in the signed permutation group. Let $(\lambda^+,\lambda^-)$ correspond to the signed permutation group conjugacy class of $w\circ (-1,1)$. Then
		$$^{L^{-1}w}\algebraicGroup{T} = \prod_{i} \restrictionOfScalars{\finiteFieldExtension{\lambda^+_i}}{\finiteField}{\multiplcativeScheme} \times \prod_{i} \restrictionOfScalars{\finiteFieldExtension{\lambda^-_i}}{\finiteField}{\algebraicGroup{\UnitaryGroup}_1} $$
	\end{enumerate}
	\item (Type C,D similitude groups)
        \begin{enumerate}[(a)]
        \item ($\algebraicGroup{GSp}$ or $\algebraicGroup{GSO}_{2n}^+$) Let $(\lambda^+,\lambda^-)$ correspond to the signed permutation group conjugacy class of $w$. Then \calvin{probably should have an alternate description}
        $$^{L^{-1}w}\algebraicGroup{T} = \{(\vec{x},\vec{y},\lambda) \in  \prod_{i} \restrictionOfScalars{\finiteFieldExtension{\lambda^+_i}}{\finiteField}{\multiplcativeScheme} \times \prod_{i} \restrictionOfScalars{\finiteFieldExtension{2\lambda^-_i}}{\finiteField}{\multiplcativeScheme}  \times \multiplcativeScheme\} | \FieldNorm{2\lambda_i^-}{\lambda_i^-}(y_i) = \lambda \textnormal{ for all }y_i\}$$
		\item ($\algebraicGroup{SO}_{2n}^-$) Let $(\lambda^+,\lambda^-)$ correspond to the signed permutation group conjugacy class of $w\circ (-1,1)$. Then
		$$^{L^{-1}w}\algebraicGroup{T} = \{(\vec{x},\vec{y},\lambda) \in  \prod_{i} \restrictionOfScalars{\finiteFieldExtension{\lambda^+_i}}{\finiteField}{\multiplcativeScheme} \times \prod_{i} \restrictionOfScalars{\finiteFieldExtension{2\lambda^-_i}}{\finiteField}{\multiplcativeScheme}  \times \multiplcativeScheme\} | \FieldNorm{2\lambda_i^-}{\lambda_i^-}(y_i) = \lambda \textnormal{ for all }y_i\}$$
        \end{enumerate}
\end{enumerate}

A summary of the types of tori are
\begin{center}
	\begin{tabular}{|c|c|c|}
		\hline 
		$\algebraicGroup{G}$ & $W(\algebraicGroup{T}^{\mathrm{std}}_{\algebraicGroup{G}})$ $\Frobenius$-conj. class& corresponding torus\tabularnewline
		\hline 
		\hline 
		$\algebraicGroup{\UnitaryGroup}$ & $\lambda = [w]$ &$\prod_{i} \begin{dcases}
	\restrictionOfScalars{\finiteFieldExtension{\lambda^+_i}}{\finiteField}{\multiplcativeScheme} & \lambda_i \textnormal{ is even} \\
	\restrictionOfScalars{\finiteFieldExtension{\lambda^+_i}}{\finiteField}{\algebraicGroup{\UnitaryGroup}_1}  & \lambda_i \textnormal{ is odd}
	\end{dcases}$  
    \tabularnewline 
    \hline
        $\begin{array}{c}
\algebraicGroup{\SO}_{2n+1}, \algebraicGroup{\Sp}_{2n}, \algebraicGroup{\SO}^{+}_{2n}\\
\algebraicGroup{\SO}^{-}_{2n}
\end{array}$ &
$(\lambda^+,\lambda^-) = \begin{cases}
[w] \\
[w\circ (-1,1)]
\end{cases}$ & $\prod_{i} \restrictionOfScalars{\finiteFieldExtension{\lambda^+_i}}{\finiteField}{\multiplcativeScheme} \times \prod_{i} \restrictionOfScalars{\finiteFieldExtension{\lambda^-_i}}{\finiteField}{\algebraicGroup{\UnitaryGroup}_1} $\tabularnewline
	\hline 
		\hline
        $\begin{array}{c}
\algebraicGroup{\GSp}_{2n}, \algebraicGroup{\GSO}^{+}_{2n}\\
\algebraicGroup{\GSO}^{-}_{2n}
\end{array}$ &
$(\lambda^+,\lambda^-) = \begin{cases}
[w] \\
[w\circ (-1,1)]
\end{cases}$ & $\prod_{i} \restrictionOfScalars{\finiteFieldExtension{\lambda^+_i}}{\finiteField}{\multiplcativeScheme} \times \left(\prod_{i} \restrictionOfScalars{\finiteFieldExtension{2\lambda^-_i}}{\finiteField}{\multiplcativeScheme}  \times_{\FieldNorm{2\lambda^-_i}{\lambda^-_i}} \multiplcativeScheme\right) $\tabularnewline
		\hline		
	\end{tabular}
\end{center}
